### 2017-10-26 paper.tex
nits

% \documentclass[10pt]{sosp2015}
% \documentclass[10pt,onecolumn]{sosp2015}
\documentclass[10pt,onecolumn,letterpaper]{article}
% \usepackage[10pt,inchmargins]{sigmin}  %% template from Xi Wang.
%\special{papersize=8.5in,11in}
%\setlength{\pdfpagewidth}{8.5in}
%\setlength{\pdfpageheight}{11in}
\usepackage[noheadfoot,
	      left=1in,right=1in,top=1in,bottom=1in,
             columnsep=0.3in
             ]{geometry}
\usepackage[small,compact]{titlesec}
\usepackage[font={small,bf}]{caption}    % added 9/10/13
\usepackage[nolineno,noindent,norules]{lgrind}
\usepackage{tightenum}
\usepackage{float}
\usepackage{xspace}
\usepackage{times,pifont}
\usepackage{mathptmx}
\usepackage{subfig,graphics,graphicx,color}
\usepackage{multirow}
\usepackage{dblfloatfix} %% correctly orders single- and double-col figures
\usepackage{hyphenat}
\usepackage{mathrsfs}
\usepackage{subfig}
\usepackage{amssymb,amsmath,centernot}
\usepackage{lastpage}
\usepackage{flushend}
\usepackage{hhline}
\usepackage{authblk}
%\newcommand{\doi}{XXXXXX}

%%% ================= START of SOSP '13 template ================= 
% \makeatletter
% 
% \def\ftype@copyrightbox{8}
% \def\@copyrightspace{
% \@float{copyrightbox}[b]
% \begin{center}
% \setlength{\unitlength}{1pc}
% \begin{picture}(20,6.0) 
% \put(0,3){\parbox{\columnwidth}{\scriptsize
% 
% %*** SAMPLE. AUTHOR PUT SUPPLIED TEXT HERE ****
% 
% \noindent
% \rule{6.0 cm}{0.2pt}\\
% Permissiondddd to make digital or hard copies of part or all of this work 
% for personal or classroom use is granted without fee provided that
% copies are not made or distributed for profit or commercial advantage 
% and that copies bear this notice and the full citation on the first
% page. Copyrights for third-party components of this work must be
% honored.  For all other uses, contact the Owner/Author. 
% 
% \vspace{\baselineskip}\noindent
% Copyright is held by the Owner/Author(s).\\
% \textit{SOSP'15}.\\
% ACM XXXXXXX.
% 
% \noindent
% http://dx.doi.org/\doi}
% }
% \end{picture}
% \end{center}
% \end@float}
% 
% \def\maketitle{\par
%  \begingroup
%    \def\thefootnote{\fnsymbol{footnote}}
%    \def\@makefnmark{\hbox
%        to 0pt{$^{\@thefnmark}$\hss}}
%      \twocolumn[\@maketitle]
% \@thanks
%  \endgroup
%  \setcounter{footnote}{0}
%  \let\maketitle\relax
%  \let\@maketitle\relax
%  \gdef\@thanks{}\gdef\@author{}\gdef\@title{}\gdef\@subtitle{}\let\thanks\relax
%  \@copyrightspace}
% 
% \makeatother

%%% ================= END of SOSP '13 template =================

%\newcommand{\comment}[1]{}
\frenchspacing

%\doublespacing

%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%     macro

\newcommand{\xxx}[0]{\textsc{Crane}\xspace}
\newcommand{\paxos}[0]{\textsc{Paxos}\xspace}
\newcommand{\mytitle}[0]{\textbf {Study Report}}
\newcommand{\mykeywords}[0]{State Machine Replication, Fault Tolerance, Stable 
and Deterministic Multithreading, Software Reliability}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
% hyperref stuff

%\usepackage[square,comma,numbers,sort]{natbib}
\usepackage{hypernat}
\usepackage{hyperref}

%% fill in pdf info here
\hypersetup{%
colorlinks=false,
pdfborder={0 0 0},
pdftitle={\mytitle},
pdfkeywords={\mykeywords},
bookmarksnumbered,
pdfstartview={FitH},
urlcolor=cyan,
pdfpagelabels=true,
pdfdisplaydoctitle=true,
}%

%\usepackage{breakurl}
%\usepackage[all]{hypcap}
%\renewcommand{\url}{\burl}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
% Some NICE fonts

\newfont{\BIG}{cminch}                             %--- One-inch font
\newfont{\sfbHuge}{cmssbx10 scaled\magstep5}       %-- 25pt sans serif bold
\newfont{\sfbLarger}{cmssbx10 scaled\magstep3}   %-- 12+pt sans serif boldd
\newfont{\sfblarger}{cmssbx10 scaled\magstep2}   %-- 12+pt sans serif bold
\newfont{\sfblarge}{cmssbx10 scaled\magstep1}      %-- 12pt sans serif bold
\newfont{\sfbeleven}{cmssbx10 scaled\magstephalf}  %-- 11pt sans serif bold
\newfont{\sfb}{cmssbx10}                           %-- 10pt sans serif bold
\newfont{\sfeight}{cmss8}                          %-- 8pt sans serif

%%%%%%%%%%%%%%%%%%%%%%%%
%    space tweaking

%\textwidth = 6.5 in
%\textheight = 9.0 in
%\setlength{\topmargin}{-.5in}

%\headheight = 0.0 in
%\headsep = 0.0 in
%\parskip = 0.2in
%\parindent = 0.0in

\renewcommand{\topfraction}{0.95}
\addtolength{\textfloatsep}{-0.1in}
%\addtolength{\floatsep}{0.025in}
\renewcommand\floatpagefraction{.9}
%\renewcommand\bottomfraction{.9}
\renewcommand\textfraction{.1}

\setlength{\parindent}{9pt}

% Rescue
\makeatletter
\def\v#1{{\mbox{\fontfamily{cmtt}\fontsize{\f@size}{\f@size}\selectfont #1}}}

\newcommand{\dmt}[0]{DMT\xspace}
\newcommand{\smt}[0]{StableMT\xspace}
\newcommand{\smr}[0]{SMR\xspace}

\newcommand{\racepro}[0]{\textsc{RacePro}\xspace}
\newcommand{\criu}[0]{\textsc{CRIU}\xspace}
\newcommand{\lxc}[0]{\textsc{LXC}\xspace}
\newcommand{\tern}[0]{\textsc{Tern}\xspace}
\newcommand{\peregrine}[0]{\textsc{Peregrine}\xspace}
\newcommand{\parrot}[0]{\textsc{Parrot}\xspace}
\newcommand{\repframe}[0]{\textsc{RepFrame}\xspace}
\newcommand{\grace}[0]{Grace\xspace}
\newcommand{\coredet}[0]{\textsc{CoreDet}\xspace}
\newcommand{\kendo}[0]{Kendo\xspace}
\newcommand{\dthreads}[0]{\textsc{DThreads}\xspace}
\newcommand{\determinator}[0]{Determinator\xspace}
\newcommand{\dos}[0]{dOS\xspace}
\newcommand{\ddos}[0]{DDOS\xspace}
\newcommand{\timealgo}[0]{time bubbling\xspace}
\newcommand{\ldpreload}[0]{LD\_PRELOAD\xspace}
\newcommand{\ntimeout}[0]{$W_{timeout}$\xspace}
\newcommand{\nclock}[0]{$N_{clock}$\xspace}

\newcommand{\apache}{\v{Apache}\xspace}
\newcommand{\mongoose}[0]{\v{Mongoose}\xspace}
\newcommand{\ab}{\v{ApacheBench}\xspace}
\newcommand{\clamav}{\v{ClamAV}\xspace}
\newcommand{\upnp}{uPnP\xspace}
\newcommand{\mediatomb}{\v{MediaTomb}\xspace}
\newcommand{\mencoder}{\v{mencoder}\xspace}
\newcommand{\mongodb}{\v{MongoDB}\xspace}
\newcommand{\ssdb}{\v{SSDB}\xspace}
\newcommand{\mysql}{\v{MySQL}\xspace}
\newcommand{\sysbench}{\v{SysBench}\xspace}
\newcommand{\zookeeper}{\v{ZooKeeper}\xspace}

\newcommand{\aget}[0]{\v{aget}\xspace}
\newcommand{\pthread}[0]{\mbox{Pthreads}\xspace}
\newcommand{\openldap}[0]{{OpenLDAP}\xspace}
\newcommand{\redis}[0]{{Redis}\xspace}
\newcommand{\bdb}[0]{{Berkeley DB}\xspace}
\newcommand{\vtune}[0]{\v{VTune}\xspace}
\newcommand{\http}[0]{\mbox{HTTP}\xspace}

% In short.
\newcommand{\eg}{{e.g.}}
\newcommand{\ie}{{i.e.}}
\newcommand{\etc}{{etc}}
\newcommand{\para}[1]{\vspace{.00in}\noindent{\bf #1}}
\newcommand{\wrt}{{w.r.t. }}
\newcommand{\cf}{{cf. }}

% Synch and network operations.
\newcommand{\checktimebubble}[0]{\v{check\_add\_timebubble()}\xspace}
\newcommand{\mutexlock}[0]{\v{pthread\_mutex\_lock()}\xspace}
\newcommand{\connect}[0]{\v{connect()}\xspace}
\newcommand{\send}[0]{\v{send()}\xspace}
\newcommand{\sendto}[0]{\v{sendto()}\xspace}
\newcommand{\sendmsg}[0]{\v{sendmsg()}\xspace}
\newcommand{\mywrite}[0]{\v{write()}\xspace}
\newcommand{\pwrite}[0]{\v{pwrite()}\xspace}
\newcommand{\close}[0]{\v{close()}\xspace}
\newcommand{\recv}[0]{\v{recv()}\xspace}
\newcommand{\select}[0]{\v{select()}\xspace}
\newcommand{\poll}[0]{\v{poll()}\xspace}
\newcommand{\epollwait}[0]{\v{epoll\_wait()}\xspace}
\newcommand{\accept}[0]{\v{accept()}\xspace}

% Parrot primitives.
\newcommand{\getturn}[0]{\v{get\_turn()}\xspace}
\newcommand{\putturn}[0]{\v{put\_turn()}\xspace}
\newcommand{\wait}[0]{\v{wait()}\xspace}
\newcommand{\signal}[0]{\v{signal()}\xspace}

% Evaluation stats.
\newcommand{\github}[0]{\url{github.com/columbia/crane}}
% \newcommand{\ntype}[0]{four\xspace}
\newcommand{\nprog}[0]{five\xspace}
\newcommand{\overheadmax}[0]{22.99\%\xspace}
\newcommand{\overhead}[0]{34.19\%\xspace}
\newcommand{\dmtspeedup}[0]{10.5\%\xspace}
\newcommand{\proxyoverhead}[0]{2.33\%\xspace}
\newcommand{\timebubblelow}[0]{66.65\%\xspace}
\newcommand{\timebubblehigh}[0]{93.88\%\xspace}
\newcommand{\recovertime}[0]{1.97ms\xspace}
\newcommand{\downgradetime}[0]{0.36s\xspace}
\newcommand{\mencoderspeedup}{\v{49\%}\xspace}

\newcommand{\us}[0]{\(\mu\text{s}\)\xspace}

\def\LGfsize{\footnotesize}
%\pagestyle{empty}

% \conferenceinfo{SOSP'15}{October 4--7, 2015, Monterey, CA}
% \copyrightyear{2015} 
% \copyrightdata{978-1-4503-3834-9/15/10} 
% \doi{2815400.2815427}

\title{\mytitle}
% \authorinfo{Authors}

% \author[+*]{Heming Cui}
% \author[*]{Rui Gu}
% \author[*]{Cheng Liu}
% \author[x]{Tianyu Chen}
% \author[*]{Junfeng Yang}
% \setlength{\affilsep}{0.5em}
% \renewcommand\AB@affilsepx{\hspace{28.0 mm}\protect\Affilfont}
% \affil[+]{\textrm\fontsize{10}{10}\selectfont The University of Hong Kong}
% \affil[*]{\textrm Columbia University}
% \affil[x]{\textrm Tsinghua University\vspace{-7.0 mm}}

\begin{document}

% Hack for: Package caption Error: No float type 'copyrightbox' defined.
%\newcounter{copyrightbox}

\date{}

\author[+]{Authors}
\maketitle
%\thispagestyle{empty}

% \begin{sloppypar}
% \begin{abstract}
% \input{abstract}
% \end{abstract}
% \end{sloppypar}

% \begin{sloppypar}
%% %\category{D.2.5}{Software Engineering}{Testing and Debugging}
%% \category{D.4.5}{Operating Systems}{Threads, Reliability}
%% \category{D.2.4}{Software Engineering}{Software/Program Verification}
%% \terms{Algorithms, Design, Reliability, Performance}
%% \keywords{\mykeywords}

%% \vskip 2mm
%% \noindent {\small \bf Categories and Subject Descriptors:} \vskip -.2mm
%% \noindent
%% {\footnotesize D.4.5~[{\bf Operating Systems}]: {Threads, Reliability}\\
%% D.2.4~[{\bf Software Engineering}]: {Software/Program Verification};}
%% \vskip 1mm
%% \noindent {\small \bf General Terms:} \vskip -.2mm
%% \noindent
%% {\footnotesize Algorithms, Design, Reliability, Performance}
%% \vskip 1mm
%% \noindent {\small \bf Keywords:} \vskip -.2mm
%% \noindent
%% {\footnotesize \mykeywords}

% \vskip 2mm
% \noindent {\small \bf Categories and Subject Descriptors:}
% {\small D.4.5~[{\bf Operating Systems}]: {Threads, Reliability};
%   D.2.4~[{\bf Software Engineering}]: {Software/Program Verification};}
% \vskip .1mm
% \noindent {\small \bf General Terms:} {\small Algorithms, Design,
%   Reliability, Performance}
% \vskip .1mm
% \noindent {\small \bf Keywords:} {\small \mykeywords}
% 
% \end{sloppypar}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
% Add page number.
\setcounter{page}{1}
\pagenumbering{arabic}

\thispagestyle{plain}
\pagestyle{plain}
\setlength{\footskip}{20pt}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

% \begin{sloppypar}
% 
% \vskip 2mm
% \noindent {\small \bf Categories and Subject Descriptors:}
% {\small D.4.5~[{\bf Operating Systems}]: {Threads, Reliability};
%   C.2.4~[{\bf Computer-communication Networks}]: {Distributed Systems};}
% \vskip .1mm
% \noindent {\small \bf General Terms:} {\small Algorithms, Design,
%   Reliability, Performance}
% \vskip .1mm
% \noindent {\small \bf Keywords:} {\small \mykeywords}
% 
% \end{sloppypar}

\begin{sloppypar}
State machine replication (\smr) is a powerful fault-tolerance 
concept~\cite{paxos:practical}. It runs replicas 
of same program on multiple nodes. To keep the replicas consistent, it invokes a 
distributed consensus protocol (typically 
\paxos~\cite{paxos:simple,lamport98parttime,paxos:complex}) to ensure that a 
quorum (typically majority) of the replicas agree on the input request sequence. 
\smr is proven to tolerate various failure scenarios, like network 
partition and packet loss.

The fault-tolerant benefit of \smr makes it particularly an attractive 
high-availability service for general server programs. Unfortunately, despite 
much effort, existing \smr systems are still hard to deploy, mainly due to two 
problems.

The first problem is performance. The consensus latency of \paxos protocols is 
notoriously high and unscalable~\cite{zookeeper,glendenning2011scalable}, 
incurring high 
performance overhead for server programs. A main reason is that messages of 
traditional TCP or UDP-based \paxos protocols have to go through software 
network layers in OS kernels. For efficiency, \paxos 
protocols typically take the Multi-Paxos approach~\cite{paxos:simple}: it 
assigns one replica as the ``leader'' to invoke consensus requests, and the 
other replicas as ``backups'' to agree on requests. To agree on an input, at 
least one round-trip time (RTT) is required between the leader and a backup. 
Given that a ping RTT in LAN typically takes hundreds of \us, and that the 
request processing time of key-value store servers (\eg, \redis) is at most 
hundreds of \us, existing \paxos protocols incur high overhead in the response 
time of server programs.

Worse, the consensus latency of extant consensus protocols is often 
\emph{scale-limited}: it increases drastically when the number of concurrent 
requests or replicas increases. For instance, 
Scatter~\cite{glendenning2011scalable} shows that the consensus latency of its 
\paxos protocol increases by 1.6X when the number of replicas increases from 3 
to 9.

\begin{figure*}[ht]
\begin{center}
\includegraphics{figures/paxos_latency}
\caption{{\em Consensus latency of four existing consensus protocols.} All four 
protocols ran a client with 24 concurrent connections. The Y axis is broken to 
fit in all protocols.}\label{fig:paxos_latency}
\end{center}
\end{figure*}

To quantify this bottleneck, we evaluated four traditional consensus 
protocols~\cite{zookeeper, spaxos:srds12, crane:sosp15, libpaxos} on 24-core 
hosts with 40Gbps network, and we spawned 24 concurrent consensus connections. 
As shown in Figure~\ref{fig:paxos_latency}, when changing the replica group 
size from 3 to 9, although network and CPUs were not saturated, the consensus 
latency of 3 protocols drastically increased by 105.4\% to 168.3\%, and 36.5\% 
to 63.7\% of the increase was in OS kernel.

When only one consensus connection 
was spawned, the latency increase on the number of replicas was more gentle.

This evaluation shows that both the number of concurrent requests and 
replicas make consensus latency increase drastically. As modern server programs 
tend to support more concurrent client connections, and advanced SMR systems 
tend to deploy more replicas (\eg, Azure~\cite{azure:book} deploys seven or 
nine replicas) to support both replica failures and upgrades, the limited 
scalability in extant consensus protocols becomes even more pronounced.

The second problem is that most existing \smr systems are not designed to 
support unmodified server programs. To utilize existing \smr services, 
developers often have to rewrite their code to orchestrate server programs into 
the narrowly defined interfaces provided by these \smr systems. For instance, 
to leverage ZooKeeper~\cite{zookeeper}, developers have to shoehorn their 
programs into the file IO interface defined by ZooKeeper.

\end{sloppypar}

% uncomment to tweak with bib spacing
%\setlength\bibsep{2.25pt}
{
%\small
 \bibliographystyle{abbrv}
 \bibliography{bib/biblio}
}

\end{document}